Utilizing Parallel Workers: \\LLVM's Vectorization Plan

### 2024-11-22 paper/paper.tex
paper

In its current state (LLVM 19.1.4) VPlan is deeply integrated into LV.

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{images/inner-loop-vplan-hcfg.png}
  \caption{Hierarchical CFG of a VPlan modelling the vectorized version of the loop in 
  Figure~\ref{fig:inner-loop-vec}.}
  \label{fig:inner-loop-vplan-hcfg}
\end{figure*}

@@

\section{Summary and Future Work}
\label{sec:summary}
Lorem ipsum

@@

>>>>

In its current state (LLVM 19.1.4) VPlan is deeply integrated into LV. To better understand the
current structure of a VPlan, modelling a certain vectorization approach, we will use the
simple conditional loop from Figure~\ref{fig:inner-loop-vec} as an example.

%TODO: Combine inner-loop-scalar-ir figure and inner-loop-vplan-hcfg into a single
%       double column figure.
\begin{figure}
  \centering
  \includegraphics[width=0.475\textwidth]{images/inner-loop-scalar-loop-ir.png}
  \caption{The scalar IR for the main loop part of the conditional loop depicted in
  Figure~\ref{fig:inner-loop-vec}. Note that \texttt{getelementptr inbounds}
  was abbreviated to \texttt{getelmptr inbs}.}
  \label{fig:inner-loop-scalar-ir}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.475\textwidth]{images/inner-loop-vplan-hcfg-loop-body-2.png}
  \caption{The vectorized loop part of a hierarchical CFG of a VPlan modelling the conditional loop
  depicted in Figure~\ref{fig:inner-loop-vec}. This VPlan has a VF of 4. Note that some keywords
  are abbreviated: \texttt{arrayidx} $\equiv$ \texttt{arridx}; 
  \texttt{getelementptr inbounds} $\equiv$ \texttt{getelmptr inbs}}
  \label{fig:inner-loop-vplan-hcfg}
\end{figure}

@@

\section{Summary and Future Work}
\label{sec:summary}
In conclusion, VPlan has been in active development for several years now. When compiling with
LLVM VPlan is used to model initial vectorization approaches, keep track of different
optimization decisions and finally execute the best VPlan to transform IR. 
However, there are still some VPlan-to-VPlan transformations and a VPlan-based cost model missing, 
before the legacy LV code can be completely discarded.

Outer loop vectorization support for VPlan is still experimental, seperated from the inner loop
LV integration, and runs on a divergent experimental path (\texttt{-enable-vplan-native-path}).
In the near future development on outer loop vectorization will continue and merging it with LV's
VPlan path is planned.

Active development of other approaches such as SLP and whole-function vectorization based on the 
VPlan architecture has not been started yet.

@@

\appendix

\section{Appendix: Additional Figures}

\begin{figure*}
  \centering
  \includegraphics[width=\textwidth]{images/inner-loop-vplan-hcfg.png}
  \caption{Full hierarchical CFG of a VPlan modelling the vectorized version of the loop 
  depicted in Figure~\ref{fig:inner-loop-vec} as of LLVM 19.1.4.}
  \label{fig:inner-loop-vplan-hcfg}
\end{figure*}
